Structured and decorated cospans

### 2019-06-16 cospans.tex
Changing some notation in Theorem 2.2 and suggesting things that we can cut out.

\node (D) at (3,0) {$d \in F(b)$};

@@

\node (A) at (0,0.5) {$a_1$};
\node (A') at (0,-0.5) {$a_1^\prime$};
\node (B) at (1,0.5) {$b$};
\node (C) at (2,0.5) {$a_2$};
\node (C') at (2,-0.5) {$a_2^\prime$};
\node (D) at (1,-0.5) {$b^\prime$};
\node (E) at (3,0.5) {$d \in F(b)$};
\node (F) at (3,-0.5) {$d^\prime \in F(b^\prime)$};

@@

(C) edge node[left]{$g$} (C')

@@

together with a morphism $\iota \colon F(h)(d) \to d^\prime$.

@@

\node (E) at (3,0) {$a_1^\prime$};
\node (F) at (4,0) {$b^\prime$};
\node (G) at (5,0) {$a_2^\prime$};
\node (H) at (4,-0.5) {$d_{M_2} \in F(b^\prime)$};
(G) edge node[above]{$o^\prime$} (F);

@@

\node (A) at (0,0) {$a_1+a_1^\prime$};
\node (B) at (1.25,0) {$b+b^\prime$};
\node (C) at (2.5,0) {$a_2+a_2^\prime$};
\node (D) at (4.25,0) {$d_{M_1 \otimes M_2} \in F(b+b')$};
(A) edge node[above]{$i+i'$} (B)
(C) edge node[above]{$o+o'$} (B);
and where the decoration on the apex is obtained using the natural transformation of the symmetric lax monoidal pseudofunctor $F:$ $$d_{M_1 \otimes M_2} \coloneqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{d_{M_1} \times d_{M_2}} F(b) \times F(b') \xrightarrow{\phi_{b,b'}} F(b+b').$$The monoidal unit 0 is given by:

@@

\node (E) at (3,0) {$a_1$};
\node (H) at (4,-0.5) {$d_M \in F(b)$};

@@

\node (A) at (0,0) {$0+a_1$};
\node (B) at (1,0) {$0+b$};
\node (C) at (2,0) {$0+a_2$};
\node (D) at (3.5,0) {$d_{0 \otimes M} \in F(0+b)$};

@@

where $d_{0\otimes M} \in F(0+b)$ is given by $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{!_0 \times d_M} F(0) \times F(b) \xrightarrow{\phi_{0,b}} F(0+b).$$The left unitor is then an isomorphism in $F\mathbb{C}\textnormal{sp}_1$ given by:
\node (A) at (0,0.5) {$0+a_1$};
\node (A') at (0,-0.5) {$a_1$};
\node (B) at (1,0.5) {$0+b$};
\node (C) at (2,0.5) {$0+a_2$};
\node (C') at (2,-0.5) {$a_2$};
\node (D) at (1,-0.5) {$b$};
\node (E) at (3.5,0.5) {$d_{0+M} \in F(0+b)$};
\node (F) at (3.5,-0.5) {$d_M \in F(b)$};

@@

(C) edge node[left]{$\ell$} (C')

@@

where $\ell$ is the left unitor of $(\mathsf{A},+,0)$, together with the isomorphism $\iota_{\lambda} \colon F(\ell)(d_{0 \otimes M}) \to d_M$. The right unitor is similar.

@@

\node (D) at (1,-0.5) {$d_{M_1} \in F(c_1)$};

@@

\node (J) at (7,0) {$c_3$};

@@

\node (L) at (7,-0.5) {$d_{M_3} \in F(c_3)$};

@@

\node (B) at (2.25,0){$(c_1+c_2)+c_3$};

@@

\node (D) at (2.25,-0.5) {$d_{(M_1 \otimes M_2) \otimes M_3} \in F((c_1+c_2)+c_3)$};

@@

where $d_{(M_1 \otimes M_2) \otimes M_3} \colon 1 \to F((c_1+c_2)+c_3)$ is given by: $$1 \xrightarrow{(d_{M_1} \times d_{M_2}) \times d_{M_3})} (F(c_1) \times F(c_2)) \times F(c_3) \xrightarrow{\phi_{c_1,c_2} \times 1} F(c_1+c_2) \times F(c_3) \xrightarrow{\phi_{c_1+c_2,c_3}} F((c_1+c_2)+c_3)$$whereas tensoring the last two and then the first results in $M_1 \otimes (M_2 \otimes M_3)$:

@@

\node (B) at (2.25,0) {$c_1+(c_2+c_3)$};

@@

\node (D) at (2.25,-0.5) {$d_{M_1 \otimes (M_2 \otimes M_3)} \in F(c_1+(c_2+c_3))$};

@@

where $d_{M_1 \otimes (M_2 \otimes M_3)} \colon 1 \to F(c_1+(c_2+c_3))$ is given by: $$1 \xrightarrow{d_{M_1} \times (d_{M_2} \times d_{M_3})} F(c_1) \times (F(c_2) \times F(c_3)) \xrightarrow{1 \times \phi_{c_2,c_3}} F(c_1) \times F(c_2+c_3) \xrightarrow{\phi_{c_1,c_2+c_3}} F(c_1+(c_2+c_3)).$$If we let $a$ denote the associator of $(\mathsf{A},+,0)$, the associator of $F\mathbb{C}\textnormal{sp}_1$ is then a map of cospans in $\mathsf{A}$ from $(M_1 \otimes M_2) \otimes M_3$ to $M_1 \otimes (M_2 \otimes M_3)$ given by:

@@

\node (B) at (2.25,0.5) {$(c_1+c_2)+c_3$};

@@

\node (D) at (2.25,-0.5) {$c_1+(c_2+c_3)$};
\node (E) at (7,0.5) {$d_{(M_1 \otimes M_2) \otimes M_3} \in F((c_1+c_2)+c_3)$};
\node (F) at (7,-0.5) {$d_{M_1 \otimes (M_2 \otimes M_3)} \in F(c_1+(c_2+c_3))$};

@@

(C) edge node[left]{$a$} (C')

@@

together with the isomorphism $\iota_a \colon F(a)(d_{(M_1 \otimes M_2) \otimes M_3}) \to d_{M_1 \otimes (M_2 \otimes M_3)}$. If we denote the above associator simply as $a$ and the left and right unitors as $\lambda$ and $\rho$, respectively, then given four objects in $F\mathbb{C}\textnormal{sp}_1$, say $M_1, M_2, M_3$ and $M_4$:

@@

\node (J) at (1,-1.5) {$c_3$};

@@

\node (L) at (1,-2) {$d_{M_3} \in F(c_3)$};

@@

\node (N) at (4,-1.5) {$c_4$};

@@

\node (P) at (4,-2) {$d_{M_4} \in F(c_4)$};

@@

then as $\mathbb{C}\textnormal{sp}(\mathsf{A})$ is a symmetric monoidal double category, the following pentagon of underlying cospans commutes:

@@

as well as the following pentagon of corresponding decorations in the category $F(c_1 +(c_2 +(c_3+c_4)))$:
\node (A) at (-.5,0) {$F(aa)(d_{((M_1 \otimes M_2) \otimes M_3) \otimes M_4}) $};
\node (B) at (2.25,1) {$F(a)(d_{(M_1 \otimes M_2) \otimes (M_3 \otimes M_4)})$};
\node (C) at (5,0) {$d_{M_1 \otimes (M_2 \otimes (M_3 \otimes M_4))}$};
\node (D) at (-.5,-1) {$F((1 \otimes a)a)(d_{(M_1 \otimes (M_2 \otimes M_3)) \otimes M_4}) $};
\node (E) at (5,-1) {$F(1 \otimes a)(d_{M_1 \otimes ((M_2 \otimes M_3) \otimes M_4)}) $};
(A) edge node[left]{$F(a)(\iota_a)$} (B)

@@

since $$F(aa)(d_{((M_1 \otimes M_2) \otimes M_3) \otimes M_4})=F((1 \otimes a)a(a \otimes 1))(d_{((M_1 \otimes M_2) \otimes M_3) \otimes M_4})$$ as the corresponding pentagon in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes.

>>>>

\node (D) at (3,0) {$x \in F(b)$};

@@

@@

@@

together with a morphism $\iota \colon F(h)(x) \to x^\prime$.

@@

\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};

@@

\node (B) at (1.25,0) {$m_1+m_2$};
\node (C) at (2.5,0) {$b_1+b_2$};
\node (D) at (4.25,0) {$x_1+x_2 \in F(m_1+m_2)$};
and where the decoration on the apex is obtained using the natural transformation of the symmetric lax monoidal pseudofunctor $F:$ $$x_1+x_2 \coloneqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2).$$The monoidal unit 0 is given by:

@@

\node (E) at (3,0) {$a$};
\node (F) at (4,0) {$m$};
\node (H) at (4,-0.5) {$x \in F(m)$};

@@

\node (A) at (0,0) {$0+a$};
\node (B) at (1,0) {$0+m$};
\node (C) at (2,0) {$0+b$};
\node (D) at (3.5,0) {$!_0 + x \in F(0+m)$};

@@

where $!_0 + x \in F(0+m)$ is given by $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{!_0 \times x} F(0) \times F(m) \xrightarrow{\phi_{0,m}} F(0+m).$$The left unitor is then an isomorphism in $F\mathbb{C}\textnormal{sp}_1$ given by:
\node (A) at (0,0.5) {$0+a$};
\node (B) at (1,0.5) {$0+m$};
\node (C) at (2,0.5) {$0+b$};
\node (C') at (2,-0.5) {$b$};
\node (D) at (1,-0.5) {$m$};
\node (E) at (3.5,0.5) {$!_0 + x \in F(0+m)$};
\node (F) at (3.5,-0.5) {$x \in F(m)$};

@@

(C) edge node[right]{$\ell$} (C')

@@

where $\ell$ is the left unitor of $(\mathsf{A},+,0)$, together with the isomorphism $\iota_{\lambda} \colon F(\ell)(!_0 + x) \to x$. The right unitor is similar.

@@

\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};

@@

@@

@@

\node (B) at (2.25,0){$(m_1+m_2)+m_3$};

@@

\node (D) at (2.25,-0.5) {$(x_1+x_2)+x_3 \in F((m_1+m_2)+m_3)$};

@@

where $(x_1+x_2)+x_3 \colon 1 \to F((m_1+m_2)+m_3)$ is given by: $$1 \xrightarrow{(x_1 \times x_2) \times x_3} (F(m_1) \times F(m_2)) \times F(m_3) \xrightarrow{\phi_{m_1,m_2} \times 1} F(m_1+m_2) \times F(m_3) \xrightarrow{\phi_{m_1+m_2,m_3}} F((m_1+m_2)+m_3)$$whereas tensoring the last two and then the first results in $M_1 \otimes (M_2 \otimes M_3)$:

@@

\node (B) at (2.25,0) {$m_1+(m_2+m_3)$};

@@

\node (D) at (2.25,-0.5) {$x_1+(x_2+x_3) \in F(m_1+(m_2+m_3))$};

@@

where $x_1+(x_2+x_3) \colon 1 \to F(m_1+(m_2+m_3))$ is given by: $$1 \xrightarrow{x_1 \times (x_2 \times x_3)} F(m_1) \times (F(m_2) \times F(m_3)) \xrightarrow{1 \times \phi_{m_2,m_3}} F(m_1) \times F(m_2+m_3) \xrightarrow{\phi_{m_1,m_2+m_3}} F(m_1+(m_2+m_3)).$$If we let $a$ denote the associator of $(\mathsf{A},+,0)$, the associator of $F\mathbb{C}\textnormal{sp}_1$ is then a map of cospans in $\mathsf{A}$ from $(M_1 \otimes M_2) \otimes M_3$ to $M_1 \otimes (M_2 \otimes M_3)$ given by:

@@

\node (B) at (2.25,0.5) {$(m_1+m_2)+m_3$};

@@

\node (D) at (2.25,-0.5) {$m_1+(m_2+m_3)$};
\node (E) at (7,0.5) {$(x_1+x_2)+x_3 \in F((m_1+m_2)+m_3)$};
\node (F) at (7,-0.5) {$x_1+(x_2+x_3) \in F(m_1+(m_2+m_3))$};

@@

(C) edge node[right]{$a$} (C')

@@

together with the isomorphism $\iota_a \colon F(a)((x_1+x_2)+x_3) \to x_1+(x_2+x_3)$. These associators and left and right unitors together satisfy the pentagon and triangle identities of a monoidal category. \textbf{Kenny: Probably cut stuff out below this.} If we denote the above associator simply as $a$ and the left and right unitors as $\lambda$ and $\rho$, respectively, then given four objects in $F\mathbb{C}\textnormal{sp}_1$, say $M_1, M_2, M_3$ and $M_4$:

@@

\node (J) at (1,-1.5) {$m_3$};

@@

\node (L) at (1,-2) {$x_3 \in F(m_3)$};

@@

\node (N) at (4,-1.5) {$m_4$};

@@

\node (P) at (4,-2) {$x_4 \in F(m_4)$};

@@

then as $\mathbb{C}\textnormal{sp}(\mathsf{A})$ is a symmetric monoidal double category, the following pentagon of underlying cospans and maps of cospans commutes:

@@

as well as the following pentagon of corresponding decorations in the category $F(m_1 +(m_2 +(m_3+m_4)))$:
\node (A) at (-.5,0) {$F(aa)(((x_1+x_2)+x_3)+x_4) $};
\node (B) at (2.25,1) {$F(a)((x_1+x_2)+(x_3+x_4))$};
\node (C) at (5,0) {$x_1+(x_2+(x_3+x_4))$};
\node (D) at (-.5,-1) {$F((1 \otimes a)a)((x_1+(x_2+x_3))+x_4) $};
\node (E) at (5,-1) {$F(1 \otimes a)(x_1+((x_2+x_3)+x_4)) $};
(A) edge node[above]{$F(a)(\iota_a)$} (B)

@@

since $$F(aa)(((x_1+x_2)+x_3)+x_4)=F((1 \otimes a)a(a \otimes 1))(((x_1+x_2)+x_3)+x_4)$$ as the corresponding pentagon in the symmetric monoidal category $(\mathsf{A},+,0)$ commutes, and then appliying the pseudofunctor $F$ to said pentagon yields a commutative pentagon in $\mathbf{Cat}$.
